# Reproducible Research: Peer Assessment 1


### Loading & preprocessing the data.
```{r loaddata}
unzip(zipfile = "repdata-data-activity.zip")
dataRead <- read.csv("activity.csv")
```


- View the loaded data frame.
```{r showdata}
names(dataRead)
str(dataRead)
head(dataRead, 10)
```

- Process/transform the data (if necessary) into a format suitable for analysis.
```{r processdata}
# Subset data frame to values without na for next process.
data_Without_NA <- dataRead[complete.cases(dataRead),]
```


### What is mean total number of steps taken per day?

- Plot a histogram of the total number of steps taken each day.
```{r}
# Find out the total steps taken per day.
totalSteps <- aggregate(steps ~ date, data_Without_NA, sum)

# Put the descriptive variable names in data frame.
names(totalSteps)[2] <- "sum_steps"

# View new created data frame.
head(totalSteps, 10)

# Plot histogram of the total steps taken per day.
hist(
        totalSteps$sum_steps,
        col = "orange",
        main = "Histogram of the (Total Number of (Steps Taken Each Day))",
        xlab = "(Total Number of (Steps Taken Each Day))",
        breaks = 30
)
```

- Calculate and report the mean and median of the total number of steps taken per day.
```{r}
mean(totalSteps$sum_steps)
median(totalSteps$sum_steps)
```


### What is the average daily activity pattern?
- Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
# the average number of steps taken, averaged across all days for each 5-minute
# Find the "interval" dataframe.
interval_dataFrame <- aggregate(steps ~ interval, data_Without_NA, mean)

# Put the descriptive variable names in data frame.
names(interval_dataFrame)[2] <- "mean_steps"

# View new created data frame.
head(interval_dataFrame, 10)

# Format plot margins to ready for long text labels.
par( mai = c(1.2, 1.5, 1,1) )

# Plot time series.
plot(
        x = interval_dataFrame$interval,
        y = interval_dataFrame$mean_steps,
        type = "l",
        main = "Time Series Plot of the 5-Minute Interval\n and the Average Number of Steps Taken, Averaged Across All Days",
        xlab = "5-Minute Interval",
        ylab = "Average Number of Steps Taken,\n Averaged Across All Days"
)
```


